\blue{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}

### 2017-05-14 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updates

Due to complex and/or heterogeneous geological formations, detailed modelling and simulation of multiphase flows through porous media is crucial to understand the aforementioned processes. Description of physics and mechanisms of multiphase flows were reported by \citet{wooding_1976} with focus on capillary pressure and the flow regimes in capillary media. A comprehensive review of mechanisms of fingering (i.e., fluid instability) and forces acting on the interface between two immiscible fluids can be found in \citet{homsy_1987}. \citet{adler_1988} described both microscopic and macroscopic scale darcy-equations \red{(BE MORE SPECIFIC ON THEIR CONTRIBUTION TO THE STATE-OF-THE-ART OF THE FIELD!!)}. Nevertheless we should refer to the area of flow instabilities  as these, first described by \citet{saffman_1958} \red{(IS THIS TRUE? SAFFMANN MAY HAVE BEEN THE FIRST TO THROUGHLY DISCUSS VISCOUS INSTABILITIES, BUT THERE WERE OTHER SCIENTISTS THAT DESCRIBED FLUID INSTABILITIES BEFORE, E.G., RAYLEIGH(1896), TAYLOR (1955) ETC)}. The first distinction should be made between microscopic and macroscopic instabilities. Microscopic instabilities describe phenomena occurring locally, for example, Kelvin-Helmholtz and Saffman-Taylor instabilities \citep{saffman_1959}. On the other hand, macroscopic instabilities involve two-phase flow system \citep{bottoni_1992}. This paper will focus on microscopic instabilities and phenomena in two-phase systems. A local flow instability that is the focus of this work is due to viscous and stress forces and often named as viscous instabilities or \textit{viscous fingering} \red{(TRY TO MERGE THESE 2 SENTENCES!)}.

@@

\medskip
Among FEM-based formulations for porous media, the control volume finite element methods~\citep[CVFEM,][]{fung_1992} has been largely used as it can guarantee local mass conservation and high-order numerical accuracy as well as being able to use tetrahedral geometry-conforming elements. In standard CVFEM formulations, pressure and velocity are interpolated using piecewise linear FE basis functions, %$\underline{\bf v}$ and $\underline{\bf p}$) 
while material properties and flow conditions (e.g., phase saturation, density, temperature, species concentration etc) are represented with CV basis functions~\citep{voller_2009}. Saturation equations are solved explicitly after solving for the dual pressure-velocity at each non-linear iteration~\citep[a detailed description of the implicit pressure explicit saturation, IMPES, formulation can be found in][]{chen_2006,lux_2012}.

@@

Since geometries are captured by finite elements, constructed control volumes typically extend on each side of the interface which may have different properties (Fig.~\ref{fig:fem_cv}a). Therefore, some average values of the coupled velocity-pressure fields (defined in the FE space but projected into the CV space) are applied across the CVs at these interfaces. These often lead to excessive numerical dispersion especially in highly heterogeneous media (represented by spatial-dependent permeability and porosity fields). In order to overcome such artificial numerical dispersion, discontinuous hybrid finite element finite volume method (DFEFVM) was introduced by \citet{nick_2011b, nick_2011a}. This novel discretisation scheme was designed to simulate flows through discrete fractured rocks in which CVs are divided along the interfaces of different materials. \citet{cumming_2011} demonstrated that CVFEM discretisation could also be used to solve Richards' equations (coupled mass conservation and Darcy equation) in heterogeneous media with relatively small computational overhead~\citep[compared with traditional coupled velocity-pressure based formulations, see also][]{cumming_phd2012}. Fluxes over CV's were calculated based upon material properties, whereas the saturation fields were volume-averaged at the interface of the materials, enforcing mass balance as described by \citet{kirkland_1992}.

@@

\medskip 
In this work, a novel CVFEM formulation, previously introduced by \citet{gomes_2017}\citep[see also][]{jackson_2015,salinas2015}, is used to conduct numerical simulations to investigate formation and evolution of viscous fingers in heterogeneous media. Continuity equation is embedded into the pressure equation to enforce mass conservation whilst ensuring force balance is preserved. A hybrid family of \PN[n]{m} triangular and tetrahedral FE pairs is used to discretise velocity and pressure, in which velocity is represented by $n$ th-order polynomials that are discontinuous across elements whereas pressure is represented by $m$ th-order polynomials that are continuous across the elements. A sketch of a \PN[1]{2} FE-pair is shown in Fig.~\ref{fig:fem_cv}a, in which velocity is represented by discontinuous and piecewise linear basis functions whereas pressure is interpolated through continuous and piecewise quadratic basis functions. Scalar fields are stored in CV space (Fig.~\ref{fig:fem_cv}b). The dual pressure and velocity fields are represented simultaneously (through non-linear projections) in FE and CV spaces.

\medskip
A brief description of the numerical formulation and viscous fluid instabilities are introduced in Sections~\ref{equations_scheme} and~\ref{section:ViscousInstabilities}, respectively. Model set up and results including initial model-benchmark are presented in Section~\ref{section:results}. Impact of mobility ratio on the fingers formation is also included in this section. The analysis considers: heterogeneity due to space-dependent permeability field, capture the development of the interface and the evolution of fingers and test-case comparisons. Finally, concluding remarks are presented in Section~\ref{Section:Conclusion}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model Formulation}\label{equations_scheme}      
%Multiphase flow in porous media is described by mass conversation and Darcy equations for all the phase $\alpha$ ($\alpha$ = g, o, w, for gas, oil, and water phases). Different domains will produce different flow instabilities. In this paper, the test-cases are based on the discontinuous Galerkin (DG) method. DG-methods combine advantages of classical finite volume and finite element methods.

@@

\medskip
The two-phase flow through a porous media domain $\Omega$, may be described by the coupled extended Darcy and saturation equations,
\begin{eqnarray}
\left(\displaystyle\frac{\mu_{\alpha}S_{\alpha}}{{\mathbf K}\mathcal{K}_{r\alpha}}\right) {\mathbf u}_{\alpha} = \underline{\underline{\sigma}}_{\alpha} {\mathbf u}_{\alpha} = -\nabla p + \mathcal{S}_{u,\alpha},\;\;\; x_{i}\in\Omega, t>0 \label{eqn:darcy_eqn} \\
\phi\displaystyle\frac{\partial S_{\alpha} }{\partial t} +   \nabla \cdot \left( {\mathbf u}_{\alpha}  S_{\alpha}\right) =  \mathcal{S}_{cty,\alpha},\;\;\; x_{i}\in\Omega, t>0\label{eqn:saturation_eqn}
\end{eqnarray}
respectively, where, $\mu$, ${\bf K}$, $p$ and $\phi$ are the viscosity, absolute permeability, pressure and porosity, respectively. ${\mathbf u}_{\alpha}$ is the saturation-weighted Darcy velocity of the $\alpha$-phase and $\mathcal{K}_{r,\alpha}$ is the relative permeability. $\mathcal{S}$ is the source term associated with the Darcy and continuity equations. $S_{\alpha}$ represents the saturation of the $\alpha$-phase with constraints of 
\begin{displaymath}
\sum\limits_{\alpha=1}^{\mathcal{N}_{p}} S_{\alpha} = 1, 
\end{displaymath}
where $\mathcal{N}_{p}$ denotes the number of phases.
%%%%%%%%%%%%%%%
%\begin{comment}
%The main equations that will be used along with the following OCVFEM techique can be summarised below. The global mass balance equation and force balance equation are solving by vanishing the velocity term and solving the system of equations for pressure. At the $n+1$ step those two equations mentioned above can be rewritten as,

@@

Finite element basis functions for velocity and pressure fields are introduced in the discretisation of force-balance equation (Eqn.~\ref{eqn:darcy_eqn}). Hybrid basis functions are also used to allow CV-based velocity to be extrapolated across the entire element. $\underline{\underline{\sigma}}_{\alpha}$ is an absorption-like term that represents the implicit linearisation of the viscous frictional forces, and lies in both CV and FEM spaces,
\begin{displaymath}
   \underline{\underline{\sigma}}_{\alpha} = \displaystyle\frac{\mu_{\alpha}S_{\alpha}}{{\mathbf K}\mathcal{K}_{r\alpha}}.
\end{displaymath}
Though, whilst saturation is calculated in CV space, permeability is piecewise constant in FE space to allow efficient representation of surface parameterised geometries (element $A$ in Fig.~\ref{fig:fem_cv}b). Saturation equations (Eqn.~\ref{eqn:saturation_eqn}) are discretised in space with CV basis function and with the $\theta$-method in time as described by \citet{gomes_book_2012}. Velocities across CV interfaces (within and between elements, see Fig.\ref{fig:fem_cv}a) are calculated through a directional-weighted flux-limited scheme based on upwind value of $\sigma$ at individual CV as described by \citet{jackson_2013}.

This formulation is applicable to $\mathcal{N}_{p}$ fluid phases and is based on two families of  FE-pairs: \PNDG[n]{m} and \PN[n]{m}~\citep{cotter_2009a}, consistent with the dual pressure-velocity representation in control volume (CV). In these families of FE-pairs velocity is represented by $n^{\text{th}}$-order polynomials that are discontinuous across elements, whereas pressure is represented by $m^{\text{th}}$-order polynomials that may be either continuous or discontinuous (thus the notation \PN[n]{m} and \PNDG[n]{m}, respectively) across elements. Mass balance (continuity) equations are solved in CV space and a Petrov-Galerkin FEM is used to obtain high-order fluxes on CV boundaries, which are limited to yield bounded fields. Simulations performed for this work were conducted using two types of elements: \PN[1]{2} and \PN[1]{1}. Saturation, and all saturation-dependent material properties such as relative permeability and capillary pressure, are represented in CV space. %Hence, each CV (and FE) is associated with a unique set of petrophysical properties, in contrast to most previous methods in which CVs span multiple values of petrophysical properties described FE-wise.

%For the simulations performed in this work, rock and fluids are assumed incompressible thus the mass balance for phase $\alpha$ can be written as
%\begin{equation}
%\phi \; \frac{\partial S_{\alpha}}{\partial t} + \nabla (u_{\alpha}S_{\alpha})= \mathcal{S}_{cty,\alpha} 
%\label{mass_balance}
%\end{equation}

@@

Equation~\ref{eqn:saturation_eqn} is discretised in time and space using the $\theta$-method. Summing over all phases yields the global mass balance equation,
\begin{eqnarray}
 && \sum_{\alpha=1}^{N_{p}} \left\{\int\limits_{\Omega_{CVi}} \; M_{i} \; \frac{\phi\left(S_{\alpha i}^{n+1}-S_{\alpha i}^{n}\right)}{\Delta t} dV\right.  + \nonumber\\
 &&  \oint_{\Gamma_{CVi}} \left[\; \theta^{n+1/2}\; {\mathbf n}\dot \; {\mathbf u}_{\alpha}^{n+1}S_{\alpha}^{n+1} \; + \; \left(1-\theta^{n+1/2}\right) \; {\mathbf u}_{\alpha}^{n}S_{\alpha}^{n}\right] d\Gamma \;- \nonumber\\
 &&  \left.\int_{\Omega_{CVi}} M_{i} \; \mathcal{S}_{cty,\alpha}^{n+\theta} \; dV\;\right\} =0,
\label{global_mass_balance}
\end{eqnarray}
where $\Omega_{CVi}$ and $\Gamma_{CVi}$ are the volume and boundary of CV i respectively, $M_{i}$ are CV basis functions, \textbf{n} is the outward pointing unit normal vector to the surface of $CV_{i}$ and $n$ is the current time level. $\theta$ varies smoothly between $0.5$ (corresponding to Crank-Nicolson method) and $1$ (corresponding to backward-Euler scheme) to avoid the introduction of spurious oscillations for large grid Courant numbers. In addition to the numerical schemes above the discretised global mass and force balance equations are solved using a multigrid-like approach, see \citet{pavlidis2016}. This numerical formulation is fully described in \citet{gomes_2017}\citep[see also][]{salinas2015,adam_2016}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Brief Overview of Viscous Instabilities}\label{section:ViscousInstabilities}
\medskip
As in many areas of fluid mechanics, the behaviour of the flow for conditions that exceed critical limits may cause instabilities in the system. The flow of a fluid over or next to another heavier fluid, can produce flow instabilities. Under this assumption when a fluid is injected in a porous domain saturated with another fluid, viscous fingering may occur, as viscous driven hydrodynamic instability leads to the collapse of the uniform interface between fluids of different viscosity.

Viscous flow instabilities can be found across many research areas, disciplines and scales, from chemical and medical separation systems to large scale geological reservoir fields. Starting from the work of \citet{muskat_1934} and later from \citet{saffman_1959} and \citet{homsy_1987}, the experimental setup of Hele-Shaw cell by \citet{saffman_1986}, and the basic mathematical model has been initially developed to describe the physics of this phenomenon\red{(SENTENCE SEEMS INCOMPLETE ... WHAT WERE THE CONTRIBUTION OF THESE AUTHORS ???)}.

\medskip
In the Hele-Shaw cell, a fluid is introduced between two parallel flat plates separated by an infinitesimally small gap. The governing equation of Hele-Shaw flows is identical to that of the inviscid potential flow and to the flow of fluid through a porous medium (Darcy's law) \red{(THIS SENTENCE IS A BIT CONFUSING ... SO THE EQNS THAT DESCRIBE FLOWS IN H-S CELLS ARE THE POTENTIAL EQUATION FOR INVISCID FLUIDS AND THE DARCY EQN? ... TRY TO REWRITE THE SENTENCE AND EXPLAIN THE IMPORTANCE OF THE H-S CELLS IN THE UNDERSTANDING OF VISCOUS FLOW INSTABILITIES)}. More recently, \citet{howison_2000} and \citet{praud_2005} gave an extensive \red{(DO YOU MEAN EXTENSIVE OR COMPREHENSIVE?)} description of the mathematical formulation of immiscible two-phase flows in Hele-Shaw cell (also know as Saffman $\&$ Taylor problem).

\medskip
 Under the assumption that fluids remain immiscible along an interface, surface tension plays an important role in determining the shape and progress of the fingers~\citep{howison_2000}. During the displacement of a fluid by a less viscous one, the expected uniform front is perturbed \citep[described by the Buckley-Leverett theory of immiscible displacement][]{buckley_1942,sheldon_1959} leading to an uneven front with elongations of the outside edge of the fluid interface (Fig.~\ref{fig:simple_case}), often referred as 'fingers'. In homogeneous domains, fingers start to develop when the surface tension acting on the interface between the fluids exert an opposite force towards the change of shape of the interface. The interface becomes unstable and collapses, taking a curved shape.

In heterogeneous domains, such instability may also be triggered by permeability differences between smaller regions as is demonstrated in \red{Section XXX}. The higher the velocity of the low viscosity fluid, the less wide (tip-splitting behaviour) the finger is. Pressure differences acting on the interface produces a net pressure force,      
\begin{equation}
\Delta P= - \gamma \; \nabla \; \cdot \; \hat{n}. 
\label{eq:pressure_dif} 
\end{equation}
This expression is also know as the Young-Laplace equation, a relation describing the capillary pressure difference across the interface between two fluids. With $\Delta P$ denoting the pressure difference, $\hat{n}$ is the unit normal vector out of the surface, and $\gamma$ is the surface tension.

As demonstrated by \citet{habermann_1960}, mobility ratio (MR, see Section~\ref{section:intro}) is a key-parameter to assess the efficiency during fluid displacement and is defined as the ratio of mobility of the displacing (fluid $i$) to that of the displaced fluid ($j$),
\begin{equation}
 \text{MR} \; = \frac{\mathcal{K}_{ri} \; \mu_{j}}{\mathcal{K}_{rj} \; \mu_{i}}. 
\label{eq:MR}
\end{equation}
MR is a function not only of fluids' viscosity and saturation, but also of the parameterised relative permeability, $\mathcal{K}_{r\alpha}$, which is often expressed as a function of the residual and maximum phase saturation prescribed in the pore rock matrix. In the simulations conducted for this work, the modified Brooks-Corey model was used \citep{alpak_1999,brooks_1964},
\begin{eqnarray}
  \mathcal{K}_{rw}\left(S_{w}\right) &=& \mathcal{K}^{\circ}_{rw}\left[\frc{S_{w}-S_{w,irr}}{1-S_{w,irr}-S_{nw,r}}\right]^{n_{w}}, \label{Eqn:CoreyBrooks1}\\
  \mathcal{K}_{rnw}\left(S_{w}\right) &=& \mathcal{K}^{\circ}_{rnw}\left[\frc{S_{nw}-S_{nw,r}}{1-S_{w,irr}-S_{nw,r}}\right]^{n_{nw}}, \label{Eqn:CoreyBrooks2}
\end{eqnarray}
where subscripts $w$ and $nw$ stand for wetting and non-wetting phases, respectively. $\mathcal{K}^{\circ}_{rw}$ and $\mathcal{K}^{\circ}_{rnw}$ are end-point relative permeability to wetting and non-wetting phases, $S_{w,irr}$ and $S_{nw,r}$ are irreducible wetting and residual non-wetting phase saturations, respectively. Exponents $n_{w}$ and $n_{nw}$ are both set to 2.

From Eqns.~\ref{eq:MR}-\ref{Eqn:CoreyBrooks2}, it is clear that during fluid displacement, the mobility ratio changes as phase saturation, $S_{w}\left(x_{i},t\right)$ and $S_{nw}\left(x_{i},t\right)$, varies in time and space (Eqn.~\ref{eqn:saturation_eqn}). Therefore, with no lack of generality, the MR can be reduced to a viscosity ratio,
\begin{displaymath}
    \text{VR} = \frc{\mu_{i}}{\mu_{j}},
\end{displaymath}
 that will be used in the parametrisation of the numerical simulations conducted in Section \ref{section:results}. Analysis performed in the following sections will make use of this simplified definition as phase saturation is a prognostic field that is calculated along with pressure and velocity. \red{(PARAGRAPH NEEDS TO BE IMPROVED)}

>>>>

Due to complex and/or heterogeneous geological formations, detailed modelling and simulation of multiphase flows through porous media is crucial to understand the aforementioned processes. Description of physics and mechanisms of multiphase flows were reported by \citet{wooding_1976} with focus on capillary pressure and the flow regimes in capillary media. A comprehensive review of mechanisms of fingering (i.e., fluid instability) and forces acting on the interface between two immiscible fluids can be found in \citet{homsy_1987}. \citet{adler_1988} described both microscopic and macroscopic scale darcy-equations. For a porous medium that is statistically homogenous two immiscible fluids are asumed to coexist together within a solid matrix. In a microscopic level fluids are newtonian and incompressible and gravity is the only external force acting on the system. An interfacial tension $\sigma$ is assumed to exist at the interfaces separating the two fluids. The main difference between microscopic and macroscopic level is the concept of capillary pressure.   
%\red{(BE MORE SPECIFIC ON THEIR CONTRIBUTION TO THE STATE-OF-THE-ART OF THE FIELD!!)}.% Nevertheless we should refer to the area of flow instabilities  as these, first described by \citet{saffman_1958} \red{(IS THIS TRUE? SAFFMANN MAY HAVE BEEN THE FIRST TO THROUGHLY DISCUSS VISCOUS INSTABILITIES, BUT THERE WERE OTHER SCIENTISTS THAT DESCRIBED FLUID INSTABILITIES BEFORE, E.G., RAYLEIGH(1896), TAYLOR (1955) ETC)}. 
The first distinction should be made between microscopic and macroscopic instabilities. Microscopic instabilities describe phenomena occurring locally, for example, Kelvin-Helmholtz and Saffman-Taylor instabilities \citep{saffman_1959}. On the other hand, macroscopic instabilities involve two-phase flow system \citep{bottoni_1992}. This paper will focus on microscopic instabilities and phenomena in two-phase systems while a this work will focus on a local flow instability due to viscous and stress forces, often referred as viscous instability or \textit{viscous fingering}.

@@

Among FEM-based formulations for porous media, the control volume finite element methods, ~\citep[CVFEM,][]{fung_1992} has been largely used as it can guarantee local mass conservation and high-order numerical accuracy as well as being able to use tetrahedral geometry-conforming elements. In standard CVFEM formulations, pressure and velocity are interpolated using piecewise linear finite element basis functions, %$\underline{\bf v}$ and $\underline{\bf p}$) 
while saturation is represented in node centered CV basis functions.  Saturation equations are solved explicitly after solving for pressure at each non-linear iteration~\citep[a detailed description of the implicit pressure explicit saturation, IMPES, formulation can be found in][]{chen_2006,lux_2012}.

@@

\medskip
%Through the $P_{1}DGP_{2}$ element type, 
Material properties and flow conditions (e.g., density, temperature, species concentration etc) can be allocated in control volume space as shown in Fig.~\ref{fig:fem_cv}. %describes the $P_{1}DGP_{2}$ element type may vary abruptly (permeability and/or porosity contrasts at fracture-matrix interfaces). 
Since geometries are captured by finite elements, constructed control volumes typically extend on each side of the interface which may have different properties. Therefore, some average value of the permeability (defined in the FE space but projected into the CV space) is applied across the CVs at these interfaces. This often leads to excessive numerical dispersion especially in highly heterogeneous media. In order to overcome such artificial numerical dispersion, discontinuous hybrid finite element finite volume method (DFEFVM) was introduced by \citet{nick_2011b, nick_2011a}. This novel discretisation scheme was designed to simulate flows through discrete fractured rocks in which CVs are divided along the interfaces of different materials. \citet{cumming_2011} demonstrated that CVFEM discretisation could also be used to solve Richards' equations (coupled mass conservation and Darcy equation) in heterogeneous media with relatively small computational overhead~\citep[compared with traditional coupled velocity-pressure based formulations, see also][]{cumming_phd2012}. Fluxes over CV's were calculated based upon material properties, whereas the saturation fields were volume-averaged at the interface of the materials, enforcing mass balance as described by \citet{kirkland_1992}.

Through the $P_{1}DGP_{2}$ element type, material properties can be described using the control volume finite element methods (Fig.\ref{fig:fem_cv}. %describes the $P_{1}DGP_{2}$ element type may vary abruptly (permeability and/or porosity contrasts at fracture-matrix interfaces). 
Since geometries are captured by finite elements, constructed control volumes typically extend over the interfaces which may have different properties. Therefore, some average value of the permeability (defined in the FE space but projected into the CV space) is applied across the CVs at these interfaces. This often leads to excessive numerical dispersion especially in highly heterogeneous media. In order to overcome such artificial numerical dispersion, discontinuous hybrid finite element finite volume method (DFEFVM) was introduced by \citet{nick_2011b, nick_2011a}. This novel discretisation scheme was designed to simulate flows through discrete fractured rocks in which CVs are divided along the interfaces of different materials. \citet{cumming_2011} demonstrated that CVFEM discretisation could also be used to solve Richards' equations (coupled mass conservation and Darcy equation) in heterogeneous media with relatively small computational overhead~\citep[compared with traditional coupled velocity-pressure based formulations, see also][]{cumming_phd2012}. Fluxes over CV's were calculated based upon material properties, whereas the saturation fields were volume-averaged at the interface of the materials, enforcing mass balance as described by \citet{kirkland_1992}.

@@

\medskip
The objective of this study is to investigate and describe the behaviour of viscous fingering through a novel CVFEM formulation that exploits adaptive mesh of fully-unstructured triangle and tetrahedron grids. Also two key characteristics that describe the research concept are: (a) the hybrid element pair $P_{n}DGP_{m}$ which provides discontinuous and piecewise representation for velocity and continuous and piecewise representation of pressure at FE space and (b) dual representation of pressure and velocity fields in the FE and CV spaces.
The objective of this study is to investigate and describe the behaviour of viscous fingering through a novel CVFE formulation that exploits adaptive mesh of fully triangle and tetrahedron grids. Also two key characteristics that describe the research concept above are: (a) the hybrid element pair $P_{n}DGP_{m}$ which provides discontinuous and piecewise representation for velocity and continuous and piecewise representation of pressure at FE space and (b) a dual representation of pressure and velocity fields in teh FE and CV spaces.

\medskip
In the next sections the numerical formulation and a summary of the main governing equations for multiphase compositional flows will discussed. The remainder of the paper will include a description of the viscous fingering along with the finite element method technique, a discussion of the test-cases and the simulation results and the key conclusions. The analysis will consider ($1$)permeability heterogeities, ($2$)capturing the development of the interface and the evolution of fingers, ($3$)test-case comparison, but for now will neglect any gravitational effects.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing equations and numerical scheme}\label{equations_scheme}      
%Multiphase flow in porous media is described by mass conversation and Darcy equations for all the phase $\alpha$ ($\alpha$ = g, o, w, for gas, oil, and water phases). Different domains will produce different flow instabilities. In this paper, the test-cases are based on the discontinuous Galerkin (DG) method. DG-methods combine advantages of classical finite volume and finite element methods.

@@

\medskip
The two-phase flow through the porous media domain $\Omega$, may be described by the coupled extended Darcy and global mass conservation equations,
\begin{eqnarray}
\left(\displaystyle\frac{\mu_{\alpha}S_{\alpha}}{{\mathbf K}\mathcal{K}_{r\alpha}}\right) {\mathbf u}_{\alpha} = \underline{\underline{\sigma}}_{\alpha} {\mathbf u}_{\alpha} = -\nabla p + \mathcal{S}_{u,\alpha},\;\;\; x_{i}\in\Omega, t>0 \label{eqn:darcy_eqn} \\
\phi\displaystyle\frac{\partial S_{\alpha} }{\partial t} +   \nabla \cdot \left( {\mathbf u}_{\alpha}  S_{\alpha}\right) =  \mathcal{S}_{cty,\alpha},\;\;\; x_{i}\in\Omega, t>0\label{eqn:saturation_eqn}
\end{eqnarray}
where, $\mu$, ${\bf K}$, $p$ and $\phi$ are the viscosity, absolute permeability, pressure and porosity, respectively. ${\mathbf u}_{\alpha}$ is the saturation-weighted Darcy velocity of the $\alpha$-phase and $\mathcal{K}_{r,\alpha}$ is the relative permeability. $\mathcal{S}$ is the source term associated with the Darcy and continuity equations. $S_{\alpha}$ is saturation of the $\alpha$-phase with the constraint of $\sum\limits_{\alpha=1}^{\mathcal{N}_{p}} S_{\alpha} = 1$, where $\mathcal{N}_{p}$ denotes the number of phases.

%%%%%%%%%%%%%%%
%\begin{comment}
%The main equations that will be used along with the following OCVFEM techique can be summarised below. The global mass balance equation and force balance equation are solving by vanishing the velocity term and solving the system of equations for pressure. At the $n+1$ step those two equations mentioned above can be rewritten as,

@@

\medskip
Finite element basis functions for velocity and pressure fields are introduced in the discretisation of force-balance equation %(Eqn.~\ref{e:darcy_eq}). 
Hybrid basis functions are also developed to allow CV-based velocity to be extrapolated across the entire element. $\underline{\underline{\sigma}}_{k}=\underline{\underline{\sigma}}_{k}\left({\mathbf K}, \mathcal{K}_{rk}, S_{K}, \mu_{k}\right)$ is an absorption-like term that lies in both CV and FEM spaces. Though, whilst the saturation field is calculated in CV space, permeability $\left({\mathbf K}\right)$ is piecewise constant in FE space to allow efficient representation of surface parametrised geometries. The global saturation equation (Eqn.~\ref{eqn:saturation_eqn}) is discretised in space with CV basis function and with the $\theta$-method in time as described by \citet{gomes_book_2012}. Velocities across CV interfaces (within and between elements, see Fig.\ref{fig:fem_cv} below) are calculated through a directional-weighted flux-limited scheme based on upwind value of $\sigma$ at individual CV as described by \citet{jackson_2013}.

\medskip
The method developed is applicable to $\mathcal{N_{p}}$ fluid phases and is based on the novel family of $P_{n}(DG)-P_{m}(DG)$ element-pair consistent with the dual pressure-velocity representation in control volume (CV) and finite element (FE) spaces. These element-pairs have a discontinuous $n^{th}$-order polynomial representation for velocity and a discontinuous $m^{th}$-order polynomial for pressure. Another element pair, that can be used is the $P_{n}DGP_{m}$ with an $m^{th}$-order continuous representation for pressure. The mass balance (continuity) equations are solved in CV space and a Petrov-Galerkin finite element (FE) method is used to obtain the high-order fluxes on CV boundaries (Fig.\ref{fig:fem_elem}), which are limited to yield bounded fields.

Saturation, and all saturation-dependent material properties such as relative permeability and capillary pressure, are represented CV-wise. The pressure, a key difference to other CVFE methods reported previously, allows the use of CVs that do not span multiple FEs. Hence each CV (and FE) is associated with a unique set of petrophysical properties, in contrast to most previous methods in which CVs span multiple values of petrophysical properties described FE-wise.

In this model we assume that rock and fluids are incompressible thus the mass ballance for phase $\alpha$ can be written as (Eq.\ref{mass_balance})

\begin{equation}
\phi \; \frac{\partial S_{\alpha}}{\partial t} + \nabla (u_{\alpha}S_{\alpha})= \mathcal{S}_{cty,\alpha}
\label{mass_balance}
\end{equation}

@@

Equation \ref{mass_balance} is discretized in time and space using the $\theta$-method %\citep{saramito_1994}. 
Summing over all phases yields the global mass balance equation %\ref{global_mass_balance},

\begin{eqnarray}
 && \sum_{\alpha=1}^{N_{p}} \int\limits_{\Omega_{CVi}} \; M_{i} \; \frac{\phi\left(S_{\alpha i}^{n+1}\right)-\left(S_{\alpha i}^{n}\right)}{\Delta t} dV  + \nonumber\\
 &&  \oint_{\Gamma_{CVi}} [\; \theta^{n+1/2}\; {\mathbf n}\dot \; {\mathbf u}_{\alpha}^{n+1}S_{\alpha}^{n+1} \; + \; (1-\theta^{n+1/2}) \; {\mathbf u}_{\alpha}^{n}S_{\alpha}^{n}] d\Gamma \;- \nonumber\\
 &&  \int_{\Omega_{CVi}} M_{i} \; \mathcal{S}_{cty,\alpha}^{n+\theta} \; dV\; =0
\label{global_mass_balance}
\end{eqnarray}

\noindent where $\Omega_{CVi}$ and $\Gamma_{CVi}$ are the volume and boundary of CV i respectively, \textbf{n} is the outward pointing unit normal vector to the surface of $CV_{i}$ and $\theta$ varies smoothly between $0.5$ (corresponding to Crank-Nicolson method) and $1$ (corresponding to backward-Euler scheme) to avoid the introduction of spurious oscillations for large grid Courant numbers. In addition to the numerical schemes above the discretised global mass and force balance equations are solved using a multigrid-like approach, see \citet{pavlidis2016}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Brief Overview of Viscous Instabilities}\label{section:ViscousInstabilities}
\medskip
As in many areas of fluid mechanics, the behaviour of the flow for conditions that exceed critical limits may cause instabilities in the system. The flow of a fluid over or next to another heavier fluid, can produce flow instabilities. Under this assumption when a fluid is injected in a porous domain saturated with another fluid trying to displace it, viscous fingering will occur, as viscous driven hydrodynamic instability leads to the collapse of the uniform interface between fluids of different viscosity. This type of instability can be found across many research areas, disciplines and scales, from small chemical and medical separation systems to large scale field geological and /or reservoir systems. Starting from the work of \citet{muskat_1934} and later from \citet{saffman_1959} and \citet{homsy_1987}, the experimental setup of Hele-Shaw cell by \citet{saffman_1986}, and the basic mathematical model has been initially developed to describe the physics of this phenomenon. The idea is to introduce a fluid between two parallel flat plates separated by an infinitesimally small gap. The governing equation of Hele-Shaw flows is identical to that of the inviscid potential flow and to the flow of fluid through a porous medium (Darcy's law). More recently work that has been done by \citet{howison_2000} and \citet{praud_2005} give an extensive description on the mathematical background of the Saffman $\&$ Taylor problem (also know as immiscible two-phase flow in Hele-Shaw cell).

\medskip
In viscous fingering, a less viscous fluid pushes a more viscous fluid, in a thin rectangular cell (known as Hele-Shaw cell). \citet{howison_2000}  Under the assumption that fluids remain immiscible along an interface, surface tension plays an important role in determining the shape and progress of the fingers. For homogeneous domains fingers will start to develop when surface tension acting on the interface, is opposing towards the change of shape of the interface thus the interface becomes unstable and will collapse, taking a curved shape. In heterogeneous environments this process may also be triggered by permeability differences between smaller regions. The higher the velocity of the injected fluid, the less wide (tip-splitting behaviour) the finger will be. In addition, pressure differences %(eq.\ref{eq:pressure_dif}) 
acting on the interface will produce a net pressure force,

\begin{equation}
\Delta P= - \gamma \; \nabla \; \dot \; \hat{n} 
\label{eq:pressure_dif} 
\end{equation}

\noindent this is also know as the Young-Laplace equation, a relation describing the capillary pressure difference across the interface between two fluids. With $\Delta P$ denoting the pressure difference, $\hat{n}$ the unit normal vector out of the surface, and $\gamma$ the surface tension. As mentioned in Section \ref{section:intro}, an unfavourable mobility-ratio in a water-oil system will have the following definition, see eqn.\ref{eq:MR}.

\begin{equation}
 MR \; = \frac{k_{rw} \; \mu_{w}}{k_{ro} \; \mu_{o}} 
\label{eq:MR}
\end{equation}

\noindent Fingering will occur in immiscible displacements when the viscocity ratio is greater than unity. %As surface tension becomes weak, the front of the steady finger is susceptible to a viscous-fingering instability by the basic mechanism associated with a less viscous fluid displacing a more viscous one. After a split, each of the new lobes of the finger is stable as a result of their being thinner than the finger from which they split. 
As surface tension becomes weak, the front is stressed and goes under a viscous-fingering instability. The front will collapse and will start splitting into new lobes of fingers. One of these fingers will eventually outgrow the others and will then spread to occupy the appropriate width of the cell. In the process, it reaches a width that is again unstable to splitting, and the pattern repeats. Thus, surface tension plays an essential dual role; it must be weak enough for the tip front to be unstable, but it is also the physical force causing the \textit{spreading} and ensuing repeated branching.

In  this paper, viscous instabilities for different test-cases will be investigated. Adaptive and unstructured mesh will be used to capture the collapse of the interface and the development of the fingers in relation to different values of mobility ratios and different maps of permeabilities. Initially test-cases that verify the existence of fingers in experimental domains will illustrated follow by more complicated computational domains and finger patterns.
